### 2016-03-23 paper.tex
Tighten prose in the Design Experiment section

\documentclass[a4paper]{article}                                     
\usepackage{url,xspace,epsfig,etoolbox}
\usepackage{natbib}
\usepackage{fullpage}
\usepackage[font=bf]{caption}
\usepackage{mdwlist}

% No page numbers
\pagestyle{empty}

% Use 'Arial' font
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

% Allow point-size specification of font
\usepackage{fix-cm}

% Allow setting margins on the abstract
\usepackage{changepage}

\newcommand{\projecttitle}{Definition and directions of cultural-narrative games}

% Prevent indentation of new paragraphs
\setlength{\parindent}{0pt}

% Put space after each paragraph. Note that this matches section titles too,
% and the workaround is below.
\setlength{\parskip}{\baselineskip}

% Allow easy overriding of section and subsection commands
\usepackage{titlesec}
\titleformat{\section}{\bfseries\fontsize{12}{14.4}\selectfont}{}{0em}{}
\titleformat{\subsection}{\bfseries\fontsize{11}{13.2}\selectfont}{}{0em}{}

% No blank line after section titles.
\titlespacing{\section}{0pt}{0pt}{-\baselineskip}
\titlespacing{\subsection}{0pt}{0pt}{-\baselineskip}

% No blank spaces between references
\setlength{\bibsep}{0pt}

% Hanging indent of 0.5 inches in references
\setlength{\bibhang}{0.5in}

% Keep all the screenshot figures the same width
\newcommand{\screenshotwidth}{2.5in}

% Is this blind or not?
\newtoggle{blind}
\toggletrue{blind}

\begin{document}

% Title is bold, centered, 14pt.
% Standard baseline skip is 1.2*fontsize, or 16.8 here.
\begin{center}
\fontsize{14}{16.8}\selectfont
\bf \projecttitle
\end{center}

% Gleaning author placement information from the existing proceedings.
\vspace{-0.25in}
\begin{center}
\iftoggle{blind}{
 % Nothing here: skip authors for blind review
}{
Paul Gestwicki, others, Ball State University\\
pvgestwicki@bsu.edu\\
}
\end{center}

\newcommand{\totan}{\textit{TotAN}}
\newcommand{\smersh}{\textit{SMERSH}}

% Abstract: 150 words or less, extra 0.5in left and right margin.
% Note that all text is justified.
%
\begin{adjustwidth}{0.5in}{0.5in}
  \textbf{Abstract:} \textit{Culture-narration games} are a promising
  genre for transformative games. This genre can be defined formal and
  dramatic characteristics, and we identify tabletop games that belong
  to the genre.  Adding technological components to the game affords
  new interaction modes as well as convenient digital distribution. We
  describe our team's pilot project at creating a technology-enhanced
  culture-narration game to improve cultural empathy among players aged
  10--14.  This pilot project highlights several complications with
  this genre, including thematic, technical, and production
  challenges.  However, the pilot project justifies continued
  experimentation, development, and formal assessment within the
  genre.
\end{adjustwidth}

\section{Introduction}

\textit{Tales of the Arabian Nights}~\citep{Goldberg2009} is a 
tabletop board game based on the eponymous folk tale.
It was originally published in 1985~\citep{Goldberg1985} but is more commonly
known for its 2009 re-release. 
It is popular within the designer board game hobby with a board game
rank of 239 (thematic rank 75) on Board Game Geek as of this writing,
and Shut Up \& Sit Down rank it as the ninth best game of all 
time~\citep{ShutUp2015}.
In this game, players control a character within a mythical Arabian
setting, and players explore the known world to accumulate Story and
Destiny points. The winner is the one who is able to meet their 
Story and Destiny point goal while having a successful encounter
in Baghdad---the City of Peace.

An intriguing property of \textit{Tales of the Arabian Nights}~(hereafter, \textit{TotAN}) is that it eschews the conventional wisdom for game design,
that player immersion is related to agency.
When a player has an encounter in \totan, they are told the name of the
encounter---such as angry merchant,  powerful prince, or
elephant's graveyard---and based only on this information,
they must choose a reaction from one of the given tables; for example, 
table~$A$ contains options to grovel, aid, rob, avoid, converse,
attack, court, abduct, or honor.
Another player then reads a corresponding entry from 
the \textit{Book of Tales}, which is a~300-page
tome containing~2600 numbered entries.
Most entries provide introductory text followed by paragraphs
tagged with skills that the active player may choose to use---without
knowing the consequences. The reader then narrates the conclusion of the
encounter and informs the player of their rewards, often in terms of
Story or Destiny points, gaining new skills, finding treasures, or
gaining status cards that modify players' options in future encounters.
To summarize, the player chooses a location on the map, but they do not 
choose what is encountered; they choose a reaction to the encounter name
without knowing how it will be interpreted in the story;
if they have the corresponding skill, they may choose to use it but
without knowing whether it will help or hinder them.
\totan{} is not a game of skill, where the player with the best tactics
wins the game: it is somewhat arbitrary and imbalanced, but from its
design emerges a fanciful gameplay experience. It is not played to win
but rather to enjoy the emerging story with friends.

This property of \totan{} reflects several characteristics
of its source material---a book that ``changed the world on a scale
unrivalled by any other literary text''~\citep[p.1]{Makdisi2008}.
Our team focused on the translation by \citep{Neil1994}, a youth-friendly
version recommended in the \totan{} rulebook.
Like the original stories, \totan{} is a collection of shorter stories
that are sometimes linked together and sometimes not.
The characters in the folk tales are often victims of fate with
very little agency over their own encounters. The games rules follow
cultural norms expressed in the folk tales:
you cannot win while sex-changed or on pilgrimage, and both marriage
and children are great blessings---unless you have an ugly baby,
which is shameful.
Baghdad is the most important city in the world, where your adventure starts
and ends.
Furthermore, many stories in \totan{} come directly from the tales,
such as the Sindbad's escape from the valley of diamonds or Aladdin's
being trapped in a magic cave.

Playing \totan{}  inspired members of our team to read translations
of the original tales, which made us recognize the cleverness of the design
and its elegant dovetailing with the source material.
Furthermore, in reading the texts, we recognized that we had already
learned elements of the stories and the culture through playing the game,
although not always consciously or explicitly.
This inspired the following analysis, in which we tease apart the
various elements of \totan{}, considering them from both game design
and learning design points of view, and then share the results of a pilot
project to create a technology-enhanced game within the same genre.

The role and meaning of narrative in games is a perennial topic
of discussion among game design theorists.
We hope to position this work not as one of ludology or narratology,
but more pragmatically as part of a systematic investigation of how
systems and stories interact to produce learning outcomes.

\section{Defining the genre of culture-narration games}

We propose that \totan{} is one of very few games comprising a genre
of \textit{culture-narration games}, and we define this genre as having
the following characteristics.

\begin{enumerate}
\item \textit{Take place in a believable, consistent setting.} 
 Although the world of \totan{} may be unfamiliar to the player,
 everything in it is representative of the world described in the original
 folk tale---complete with the contradictions that add depth and nuance.

\item \textit{Use narrative as a primary feedback mechanism.} 
 In \totan{}, the player's reaction and skill choice yields two forms
 of feedback: the narrative description followed by the changes to the
 game state. The narrative is primary, both chronologically and aesthetically.
 Note that we follow \citet{Koster2012} in treating narrative as a 
 feedback mechanism, not a game mechanism.

\item \textit{Have measurable goals.} There is a winning condition as part
 of the social contract of play, following colloquially-accepted 
 standards for board games. The ``game'' is not simply constrained
 cooperative storytelling or a role-playing experience without formal end
 conditions.

\item \textit{Incorporate endogenously meaningful ambiguous decisions.}
 Following \citet{Burgun2012}, the decisions that one makes in the game
 are meaningful even though they are made without complete knowledge
 of the game state. For example, a player makes 
 the choice of reaction in \totan{} in hopes that it leads toward skills
 that they have, such as choosing ``Fight'' while in possession of the
 ``Weapon Use'' skill, even though one does not know whether or not
 this skill will have relevance in the resulting entry from the
 \textit{Book of Tales}.

\item \textit{Reward players for decision-making that reflects
    cultural understanding.}  The knowledge that a player brings to
  bear on decisions is not just knowledge of in-game systems.  If one
  chooses to Fight an Angry Ifrit in \totan{} without the Weapon Use
  skill, it is probably not going to end well. This prediction is not
  based on my having memorized the \textit{Book of Tales}, nor knowing how many
  hit dice an Ifrit has (as in \textit{Dungeons \& Dragons} powergaming).
  Rather, this prediction draws upon a cultural understanding of anger,
  spirits, violence, and likely much more.
  However, the game does this without establishing
  ``right'' and ``wrong'' (``moral/immoral'',
  ``light side/dark side'', ``paragon/renegade'') choices.  There
  remain unexpected twists: perhaps fighting the Angry Ifrit without
  combat skills makes him respect you and grant a boon.  \totan{},
  after all, is a game that allows the player to choose to ``drink'' a
  violent storm or ``enter'' a mysterious artifact.  While the world
  remains consistent, it also embraces the potential for surprising
  results.

\end{enumerate}

Two formal elements from \totan{} are intentionally not included in the
genre definition above. Perhaps the most clear omission is players' reading
each others' stories. While we agree that this is a critical element of 
\totan{}---and that it dovetails performatively 
and practomimetically~\citep{Travis2011} with Scheherazade's
performances in the source material---it is not clear that this is
characteristic to the genre.  Anecdotal evidence shows that
players with low reading skills can ruin the play experience for other
players, while those players could have read their own stories, aloud or not.
Similarly, \totan{} permits a single-player experience which, while
having a different aesthetic, still seems to produce the same general outcomes
we described above.

The other formal element omitted from the genre definition is the use of an
explicit map. The map is an important tangible aspect of \totan{}, 
in particular with its representation of Baghdad as the largest, most
important city in the world. However, we also note that virtual spaces
can be represented without such an explicit map, such as in the MUD family
of games or choose-your-own-adventure books, both of which
use discrete spaces without giving them visual manifestation.

Analyzing \textit{Agents of SMERSH}~\citep{Maxwell2012} establishes that
these characteristics form a genre and not merely selected elements from one
game.
\smersh{} is a cooperative game set in the Cold War, with players as 
covert agents trying to stop the enigmatic Dr.\ Lobo. 
Players move across the world collecting resources and battling henchmen,
facing encounters clearly inspired by \totan{}: flip a card to read an
encounter title, choose a reaction from a table, hear an introduction,
and then get different results based on skills. Unlike \totan{},
\smersh{} uses dice for story resolution, and supporting resource
management mechanisms introduce important strategic decisions.
Despite the use of dice for story resolution, the feedback is still
primarily narrative.

For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game about building a city both above and below the
surface.  \textit{Above and Below} is primarily about building an
economic engine that generates victory points faster than your
opponents within a fixed number of rounds. Exploring the underground
is one source of points, and this is done by determining a random
entry within its encounter book.  A player reads this entry aloud,
after which the player is presented with a decision of how to
react. However, the reaction titles are a thin wrapper around a
risk-and-reward balance: each has a number which is the number of
successes required on dice rolls, so players choose the highest number
they believe they can roll. Higher numbers always equate to better results,
and this roll is usually the end of the encounter: a successful player
is given an in-game reward with no other story or narrative.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
to be in the culture-narration game genre.

\section{Cultural Empathy}

We have discussed how \totan{} uses cultural understanding, not only on a
small scale as a way to reward players, but also as a key takeaway of
the game. In order for the game to reward players for their cultural
understanding, we have to present cultures in a way that they can, to
some extent, be understood. In other words, our game needs to foster a
sense of \textit{cultural empathy}, defining ``empathy'' as 
the ``state of mind in which someone shares the feelings or outlook of
another, sometimes prompted by imaginative exercises 'stepping into
someone's shoes'''~\citep[p.242]{Honderich2005}
Therefore, cultural
empathy, loosely formed, is having empathy for others who are from
different cultures, which allows for better understanding of that
culture.

This goal of cultural empathy differs from a goal of \textit{cultural
literacy} which prepares individuals to participate in another culture
based on the information of the culture that they have been 
given~\citep[p.165]{Hirsch1983}. Being culturally literate is having the
ability to be able to participate in another culture that you are
``dropped into,'' similar to how being literate is having the ability to
understand or write a given work of text.  \totan{} focuses on a certain
level of literacy of the given culture. Players are rewarded for their
cultural understanding and their ability to ``perform'' in the given
culture because of the information they have. This is demonstrated in
the example above, understanding the cultural views of marriage and
sex-change have an important role in game play. However, cultural
empathy makes the main point the understanding, not the
performance. \totan{} relies on an understanding of cultural values as
opposed to just focusing on some of their values consequential
norms. Therefore, though \totan{} expects some sense of cultural
literacy, it focuses more on cultural empathy. When a player walks
away from the game they would not have gained the ability to
participate in either a Middle Eastern or South Asian culture;
however, they will have gained an understanding of that culture. 
%This
%is why our emphasis in creating the game has relied more on cultural
%empathy than cultural literacy. We focus more on a mindset of a
%culture, than about facts about the culture. 
In the context of the
game, cultural empathy is a means to the end of cultural
literacy. However, outside of the game, focusing on cultural literacy
within a semi-fictional world works as the means for cultivating
cultural empathy, particularly world culture empathy.

\section{Context}

Investigating the potential for culture-narration games to teach
cultural empathy brought our team into a partnership with
\textit{a local library branch hidden for blind review}. This branch
serves primarily patrons from low socio-economic status, although
its special technology education programs also draw homeschoolers from
around the region.
The youth patrons of the library tend to be 4rd--6th grade 
``latchkey kids'' who 
see the library as a place to visit with friends, to participate in programs,
and to access technology.

Conversations with program coordinators about culture-narration games
revealed that despite other differences, the youth patrons shared
a fascination with monsters, horror, and the macabre. 
This is in part because of cultural taboo: the youth will brag about
being able to see a PG-13-rated movie or play an M-rated game,
for example, and several kids worked together to produce an original
short movie based on battles between movie monsters.

The fascination exhibited by the youth patrons reflects various theses
of Cohen's monster theory~\citep{Cohen1996}. 
We see in their behavior that ``the same creatures who terrify and
interdict can evoke potent escapist fantasies''~(p.16).
Recognizing the youth patrons' attraction to this theme,
Cohen's theory establishing monsters as reflections of culture,
and the narrative potential of monsters,
our team decided to engage in a design experiment to validate our
hypotheses about culture-narration games,
in collaboration with \textit{the library branch}.

\section{Design Experiment}

Our experiment involved a team of eleven undergraduate students and a
faculty mentor.  The students majors include Computer Science, English
(Creative Writing), Philosophy, Telecommunications (Audio Production),
and Art (Animation).
Each student earned three credit hours for the experience during a fifteen-week
semester,
working together in a studio space for two hours each Monday, Wednesday,
and Friday.
The team collectively agreed upon three primary goals for the design
experiment in the cultural-narration genre:
to use narrative as the primary form of player feedback;
to teach cultural empathy;
and to use technology to address complications of analog implementations.

After these goals were defined, the team figured out a game concept, and 
developed it throughout three iterations.
The deadlines for each iteration provided the team a chance to reflect on both
the project itself and on our methodology.  Each iteration was tested at
\textit{the library branch}, providing valuable data for which directions we
should bring the project.  We also had reflective meetings in which we 
discussed what worked about our methodology, and what should change.  During
the first two iterations, we organized our tasks using \textit{story mapping}.
However, we realized during our second reflective meeting that it might be
better to switch to Scrum.

We used JavaFX for rough prototyping, which some of the developers had used
before, and then moved to \textit{PlayN}, a cross-platform library that allowed
us to build both for Java and HTML.
Our writers and artist continually developed and improved assets for the game,
learning how to bring them into the game easily.  The writers also developed a
methodology for workshopping each other to make sure the narrative was 
developed and had a consistent voice.

Most of the team played vital roles in both the research side of the game and
the development side of the game.  The writers, for example, did extensive
research on different cultures and monsters as well as writing in a way that
would be easy for kids to read.  Writers and developers were involved in the
playtesting as well, and gained valuable insight on what the kids needed and
wanted.

We also invited guests to join us during our regular meeting times to talk about
their experiences with games, monsters, and children's literacy.  One of our
guests gave us valuable insight on what monsters mean to people and how they are
influenced by culture.

\section{Pilot Study}

\textit{In which we describe what playtesting has revealed to us}

\section{Conclusions and Future Work}

Based on our design experiments, pilot study, and understanding of the
literature, we believe there is a significant potential for
culture-narration games as learning tools, particularly for
cultural empathy.
However, as with any educational intervention, these games do not
come without complications.

Physical games such as \totan{} and \smersh{} are expensive to produce
and, for many, intimidating to encounter. \totan{}'s \textbf{X}-page
instruction booklet may not faze a board game enthusiast, but we have
observed that others are intimidated by the large rulebook, heavy box,
and many components included in the game. Combined with the duration
of play and complexity of rules, such games seem inappropriate for use
in almost all formal educational settings such as conventional
schools.  However, we have found that the combination of technology
enhancement, grade-level writing, and shorter duration circumvents
these problems.  Digital implementations all but eliminate
distribution costs, and with clever game design, children can play the
game in the relatively chaotic environment of an informal after-school
program. With curriculum aides and improved scaffolding, such as
post-game debriefing, we are certain the game could be easily
integrated into school environments, particularly those with
ready access to computers.

The role of text within the game is worthy of particular attention.
The textual expression of narrative requires that players already possess
literacy skills: players who struggle with vocabulary or decoding words
are easily frustrated by such games. There is an opportunity
here as well, to differentiate text based on player reading level
or provide embedded reading aids; however, this also transforms
the outcome of the game from cultural empathy to reading literacy.
Player oration---having players read aloud to each
other---additionally requires oratory skills. We observed engagement
being enhanced when the reader employs theatrical talents as well.
The precise role of oratory skill, self-confidence, and the social
contract of gameplay (the ``magic circle'') is an area for future work.

\section{Acknowledgments}
\textit{The team, organizational, and community partner
  acknowledgments are hidden for blind review.}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}

### 2016-03-23 paper.tex
Specify rulebook length

@@
 Physical games such as \totan{} and \smersh{} are expensive to produce
-and, for many, intimidating to encounter. \totan{}'s \textbf{X}-page
+and, for many, intimidating to encounter. \totan{}'s 19-page
 instruction booklet may not faze a board game enthusiast, but we have

### 2016-03-23 paper.tex
Revise design experiment section including several new references

@@
 and Friday.
+% In future version, cite the TOCE paper
 The team collectively agreed upon three primary goals for the design
@@
 
-After these goals were defined, the team figured out a game concept, and 
-developed it throughout three iterations.
-The deadlines for each iteration provided the team a chance to reflect on both
-the project itself and on our methodology.  Each iteration was tested at
-\textit{the library branch}, providing valuable data for which directions we
-should bring the project.  We also had reflective meetings in which we 
-discussed what worked about our methodology, and what should change.  During
-the first two iterations, we organized our tasks using \textit{story mapping}.
-However, we realized during our second reflective meeting that it might be
-better to switch to Scrum.
-
-We used JavaFX for rough prototyping, which some of the developers had used
-before, and then moved to \textit{PlayN}, a cross-platform library that allowed
-us to build both for Java and HTML.
-Our writers and artist continually developed and improved assets for the game,
-learning how to bring them into the game easily.  The writers also developed a
-methodology for workshopping each other to make sure the narrative was 
-developed and had a consistent voice.
-
-Most of the team played vital roles in both the research side of the game and
-the development side of the game.  The writers, for example, did extensive
-research on different cultures and monsters as well as writing in a way that
-would be easy for kids to read.  Writers and developers were involved in the
-playtesting as well, and gained valuable insight on what the kids needed and
-wanted.
-
-We also invited guests to join us during our regular meeting times to talk about
-their experiences with games, monsters, and children's literacy.  One of our
-guests gave us valuable insight on what monsters mean to people and how they are
-influenced by culture.
+Given these goals, the team proceeded with an incremental and
+iterative approach to game design and development, each iteration
+producing an executable release in accordance with best practices of
+agile software development~\citep[see, for
+example,][]{Cockburn2006,Keith2010}.  Each iteration was tested at
+\textit{[the library branch]}, providing valuable feedback to the
+team.  Each iteration concluded with a retrospective
+meeting~\citep{Kerth2001}, which provided an opportunity for the
+team's reflective practice~\citep{Schon1984}.
+The results of these retrospectives included large and small changes
+to the team's methodology, including a transition from
+story mapping~\citep{Patton2014} to Scrum~\citep{Schwaber2013}.
+In addition to paper prototypes, the team 
+developed prototypes using JavaFX, PlayN, and Polymer, 
+all of which allow for HTML5 Web deployment.
+Game images and stories were inspired by the team's research on monsters
+from world cultures, which included both research sources
+(such as \citet{Cohen1996}) and children's books (especially
+\citet{Steer2008} and \citet{Stowell2013}).
+The team also consulted with local university experts on
+games, monsters, world cultures, and literacy.
 

### 2016-03-23 paper.tex
Add three images and mention design logs

@@
 \usepackage{mdwlist}
+\usepackage{graphicx}
 
@@
 
-Our experiment involved a team of eleven undergraduate students and a
-faculty mentor.  The students majors include Computer Science, English
+A team of eleven undergraduate students, working with a faculty mentor,
+undertook a design experiment and pilot study to explore a monster-themed
+culture-narration game.
+The students' majors include Computer Science, English
 (Creative Writing), Philosophy, Telecommunications (Audio Production),
@@
 Given these goals, the team proceeded with an incremental and
-iterative approach to game design and development, each iteration
-producing an executable release in accordance with best practices of
-agile software development~\citep[see, for
-example,][]{Cockburn2006,Keith2010}.  Each iteration was tested at
+iterative approach to game design and development.
+The team used a design log to record design decisions~\citep{Cook2011},
+and each iteration
+produced an executable release in accordance with best practices of
+agile software development~\citep[see, for example,][]{Cockburn2006,Keith2010}. 
+Figures~\ref{fig:screenshots} and~\ref{fig:paper-prototype} demonstrate
+design artifacts from the design and development process.
+Each iteration was tested at
 \textit{[the library branch]}, providing valuable feedback to the
@@
 games, monsters, world cultures, and literacy.
+
+\newcommand{\figwidth}{2.5in}
+\begin{figure}
+\centering
+\includegraphics[width=\figwidth]{iteration-1}
+\includegraphics[width=\figwidth]{kappa-encounter}
+\caption{Screenshots from Iteration 1 (left) and 2 (right)}
+\label{fig:screenshots}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\figwidth]{paper-prototype}
+\caption{Paper prototype from Iteration 3}
+\label{fig:paper-prototype}
+\end{figure}
 

### 2016-03-23 paper.tex
Change title
Add citation to constructionism to help justify our approach

@@
 
-\newcommand{\projecttitle}{Definition and directions of cultural-narrative games}
+\newcommand{\projecttitle}{The Culture-Narration Genre: Defining the genre and identifying its potential for educational games}
 
@@
 cultural empathy brought our team into a partnership with
-\textit{a local library branch hidden for blind review}. This branch
+\textit{[a local library branch hidden for blind review]}. This branch
 serves primarily patrons from low socio-economic status, although
@@
 undertook a design experiment and pilot study to explore a monster-themed
-culture-narration game.
+culture-narration game;
+our approach draws upon constructionist~\citep{Papert2001}
+traditions, engaging in the design experiment with a multifaceted approach,
+building prototypes as a way to understand them.
 The students' majors include Computer Science, English

### 2016-03-24 paper.tex
Use microtype package for improved typography
Interleaved Kay's monster theme section into the context section

@@
 \documentclass[a4paper]{article}
+\usepackage{microtype}
 \usepackage{url,xspace,epsfig,etoolbox}
@@
 
-\section{Context}
+\section{Context and Theme}
 
-Investigating the potential for culture-narration games to teach
-cultural empathy brought our team into a partnership with
-\textit{[a local library branch hidden for blind review]}. This branch
-serves primarily patrons from low socio-economic status, although
-its special technology education programs also draw homeschoolers from
-around the region.
+Cultivating world culture empathy in children through narrative driven
+gameplay begs for a theme that is inviting on multiple accounts. We
+were drawn to the question of what our target audience looked for in
+narration, in addition to what would serve as an aid in encouraging
+world cultural empathy. 
+We formed a partnership with \textit{[omitted for blind review]}, 
+a non-lending branch that 
+provides acommunity, training, and technology services in a
+low socioeconomic area of our city.
 The youth patrons of the library tend to be 4rd--6th grade 
@@
 Conversations with program coordinators about culture-narration games
-revealed that despite other differences, the youth patrons shared
-a fascination with monsters, horror, and the macabre. 
-This is in part because of cultural taboo: the youth will brag about
-being able to see a PG-13-rated movie or play an M-rated game,
-for example, and several kids worked together to produce an original
-short movie based on battles between movie monsters.
+revealed that despite other differences, the youth patrons shared a
+fascination with monsters, horror, and the macabre.  This answer in
+some ways seemed unsatisfactory: horror and empathy---cultural or
+otherwise---do not seem to be a good match. Indeed, enjoying horror, when
+viewed simply as enjoyment of others' pain, seems like the antithesis
+of empathy. However, this love of horror, or the monstrous, can be
+seen in our culture, touching down on all forms of media. Monsters has
+been a hook for narratives throughout history. They have attracted
+people of various walks of life, by playing into the fears of the
+people.
 
-The fascination exhibited by the youth patrons reflects various theses
-of Cohen's monster theory~\citep{Cohen1996}. 
-We see in their behavior that ``the same creatures who terrify and
-interdict can evoke potent escapist fantasies''~(p.16).
-Recognizing the youth patrons' attraction to this theme,
-Cohen's theory establishing monsters as reflections of culture,
-and the narrative potential of monsters,
-our team decided to engage in a design experiment to validate our
-hypotheses about culture-narration games,
-in collaboration with \textit{the library branch}.
+\citeauthor{Cohen1996} draws attention to this in his book, 
+\textit{Monster Theory}~\citeyear{Cohen1996}.
+He offers a way of ``reading cultures through the monsters they
+engender'' (p.~3).  Cohen notes that the conclusions we come to about
+the cultures, from monsters or otherwise, will be a combination of
+many fragments rather than a ``smooth epistemological whole.'' A monster
+is brought about at a particular time and represents not only the
+fears prevalent in the culture but also the values, that lead to that
+fear (p.~4--12). Pinning that down is difficult. The monsters work more
+as a ``snapshot'' of the culture at a particular moment than a
+definition. This point is especially important for us to note when
+creating a game whose goal is to encourage children's efforts to see
+something from others' perspectives.
+%Yet, we see in the children's behavior that ``the same creatures who terrify and interdict can evoke potent escapist fantasies''~(p.16).
+
+For these reasons, monsters have become a staple theme for our
+project. They were not only enticing to our target
+audience---first observed by our community partner and later proven
+through play testing---but also provide an excellent passage for
+encouraging empathy, particularly cultural empathy. They represent
+various cultures' sameness, in base forms of fear, yet difference, in
+the ways that fear becomes specific and embodied. By encouraging an
+understanding of both difference and sameness between cultures, they
+encourage individuals to try to see something from someone else's
+viewpoint.
 

### 2016-03-24 paper.tex
Finished complete read-through
This version is 8 pages, which is too long.

--- 2016-03-23 paper.tex
@@
 \usepackage{graphicx}
+\usepackage{wrapfig}
 
@@
 
-% Keep all the screenshot figures the same width
-\newcommand{\screenshotwidth}{2.5in}
+% Allow for bigger figures not on their own pages.
+% http://tex.stackexchange.com/questions/68516/avoid-that-figure-gets-its-own-page
+\renewcommand{\floatpagefraction}{.8}
 
@@
 }{
-Paul Gestwicki, others, Ball State University\\
-pvgestwicki@bsu.edu\\
+Paul Gestwicki, Kaley Rittichier, and Austin DeArmond, Ball State University\\
+pvgestwicki@bsu.edu, kjrittichier@bsu.edu, ajdearmond@bsu.edu\\
 }
@@
 \textit{Tales of the Arabian Nights}~\citep{Goldberg2009} is a 
-tabletop board game based on the eponymous folk tale.
-It was originally published in 1985~\citep{Goldberg1985} but is more commonly
-known for its 2009 re-release. 
-It is popular within the designer board game hobby with a board game
-rank of 239 (thematic rank 75) on Board Game Geek as of this writing,
+tabletop board game based on the eponymous folk tale,
+originally published in 1985~\citep{Goldberg1985} and
+released with many revisions in 2009.
+It is popular among board game hobbyists, having a board game
+rank of~239 (thematic rank~75) on Board Game Geek as of this writing,
 and Shut Up \& Sit Down rank it as the ninth best game of all 
@@
 In this game, players control a character within a mythical Arabian
-setting, and players explore the known world to accumulate Story and
+setting, exploring the known world to accumulate Story and
 Destiny points. The winner is the one who is able to meet their 
 Story and Destiny point goal while having a successful encounter
-in Baghdad---the City of Peace.
+in Baghdad---the City of Peace. Notice in Figure~\ref{fig:totan-board}
+that Baghdad is the center of the world, much larger and more beautiful
+than any other city.
+
+\newcommand{\figwidth}{2.5in}
+%\begin{wrapfigure}{R}{\figwidth}
+\begin{figure}[b]
+\centering
+\includegraphics[width=\figwidth]{totan-board}
+\caption{Center of the \textit{Tales of the Arabian Nights} game board~\citep{Goldberg2009}.}
+\label{fig:totan-board}
+%\end{wrapfigure}
+\end{figure}
 
 An intriguing property of \textit{Tales of the Arabian Nights}~(hereafter, \textit{TotAN}) is that it eschews the conventional wisdom for game design,
-that player immersion is related to agency.
-When a player has an encounter in \totan, they are told the name of the
+that player immersion is dependent upon agency.
+In \totan{}, 
+players choose a destination without any foreknowledge about
+what might be encountered there.
+They are told the name of the
 encounter---such as angry merchant,  powerful prince, or
 elephant's graveyard---and based only on this information,
-they must choose a reaction from one of the given tables; for example, 
-table~$A$ contains options to grovel, aid, rob, avoid, converse,
+they must choose a reaction from a set of options; for example, 
+table~$A$ permits a choice of grovel, aid, rob, avoid, converse,
 attack, court, abduct, or honor.
 Another player then reads a corresponding entry from 
-the \textit{Book of Tales}, which is a~300-page
-tome containing~2600 numbered entries.
+the \textit{Book of Tales}---a~300-page tome containing~2600 numbered entries.
 Most entries provide introductory text followed by paragraphs
-tagged with skills that the active player may choose to use---without
-knowing the consequences. The reader then narrates the conclusion of the
-encounter and informs the player of their rewards, often in terms of
-Story or Destiny points, gaining new skills, finding treasures, or
+tagged with skills that the active player may choose to use, but the choice
+is made without any knowledge of potential consequences.
+The reader then narrates the conclusion of the
+encounter and informs the player of their rewards, which might include
+gaining Story or Destiny points, learning new skills, finding treasures, or
 gaining status cards that modify players' options in future encounters.
-To summarize, the player chooses a location on the map, but they do not 
-choose what is encountered; they choose a reaction to the encounter name
-without knowing how it will be interpreted in the story;
-if they have the corresponding skill, they may choose to use it but
-without knowing whether it will help or hinder them.
-\totan{} is not a game of skill, where the player with the best tactics
-wins the game: it is somewhat arbitrary and imbalanced, but from its
-design emerges a fanciful gameplay experience. It is not played to win
-but rather to enjoy the emerging story with friends.
+\totan{} then is hardly a game of skill, where the player with the best tactics
+wins the game: viewed through this conventional lens, the game is arbitrary
+and imbalanced. However, from its design emerges a fanciful gameplay
+experience.
+Although it is a competitive board game, 
+it is not played to win,
+but rather for the enjoyment of emerging narrative,
+similar to tabletop roleplaying games~\citep{Grouling2001}
+or story games~\citep{Duncan2015}.
 
@@
 unrivalled by any other literary text''~\citep[p.1]{Makdisi2008}.
-Our team focused on the translation by \citep{Neil1994}, a youth-friendly
-version recommended in the \totan{} rulebook.
 Like the original stories, \totan{} is a collection of shorter stories
@@
 Playing \totan{}  inspired members of our team to read translations
-of the original tales, which made us recognize the cleverness of the design
+of the original tales,
+particularly  the youth-friendly translation by \citep{Neil1994},
+which made us recognize the cleverness of the design
 and its elegant dovetailing with the source material.
+Indeed, although the game was not explicitly designed as a learning game,
+we find in it exhibits many of the learning game design principles
+established by \citet{Klopfer2009}.
 Furthermore, in reading the texts, we recognized that we had already
@@
 
-The role and meaning of narrative in games is a perennial topic
-of discussion among game design theorists.
-We hope to position this work not as one of ludology or narratology,
-but more pragmatically as part of a systematic investigation of how
-systems and stories interact to produce learning outcomes.
+\section{Defining the genre}
 
-\section{Defining the genre of culture-narration games}
-
-We propose that \totan{} is one of very few games comprising a genre
-of \textit{culture-narration games}, and we define this genre as having
-the following characteristics.
+We propose that \totan{} is one of very few published games comprising a genre
+of \textit{culture-narration games}. Recognizing that the term ``genre''
+can be interpreted in multiple ways, we adopt the perspective established
+by \citet{Arsenault2009},
+that a genre is ``the codified usage of particular mechanics
+and game design patterns to express a range of intended play-experiences.''
+We define this genre, then, as having the following characteristics.
 
@@
  everything in it is representative of the world described in the original
- folk tale---complete with the contradictions that add depth and nuance.
+ folk tale.
+ This includes the contradictions that add depth and nuance, such as 
+ the inconsistent treatment of outsiders that emerged from the 
+ complex historical roots of the folk tales themselves~\citep{Matar2008}.
 
@@
  of feedback: the narrative description followed by the changes to the
- game state. The narrative is primary, both chronologically and aesthetically.
- Note that we follow \citet{Koster2012} in treating narrative as a 
- feedback mechanism, not a game mechanism.
+ game state. The narrative is primary---temporally, visually, and aesthetically.
 
@@
 
-\item \textit{Incorporate endogenously meaningful ambiguous decisions.}
+\item \textit{Incorporate endogenously-meaningful ambiguous decisions.}
  Following \citet{Burgun2012}, the decisions that one makes in the game
@@
 
-\item \textit{Reward players for decision-making that reflects
+\item \textit{Reward players for decisions that reflects
     cultural understanding.}  The knowledge that a player brings to
-  bear on decisions is not just knowledge of in-game systems.  If one
-  chooses to Fight an Angry Ifrit in \totan{} without the Weapon Use
-  skill, it is probably not going to end well. This prediction is not
-  based on my having memorized the \textit{Book of Tales}, nor knowing how many
-  hit dice an Ifrit has (as in \textit{Dungeons \& Dragons} powergaming).
-  Rather, this prediction draws upon a cultural understanding of anger,
-  spirits, violence, and likely much more.
-  However, the game does this without establishing
-  ``right'' and ``wrong'' (``moral/immoral'',
+  bear on decisions is not just knowledge of in-game systems. 
+  A player can predict that choosing to fight a Powerful Ifrit 
+ without the Weapon Use skill will not end well,
+  but this prediction is not made by having memorized the 
+  \textit{Book of Tales}.
+  Rather, this prediction draws upon a cultural understanding of power,
+  spirits, violence, and much more.
+  Note that the game rewards player decisions without resorting to
+  clear-cut ``right'' and ``wrong'' (``moral/immoral'',
   ``light side/dark side'', ``paragon/renegade'') choices.  There
-  remain unexpected twists: perhaps fighting the Angry Ifrit without
+  remain unexpected twists: perhaps fighting the Powerful Ifrit without
   combat skills makes him respect you and grant a boon.  \totan{},
   after all, is a game that allows the player to choose to ``drink'' a
-  violent storm or ``enter'' a mysterious artifact.  While the world
+  Violent Storm or ``enter'' a Mysterious Artifact.  While the world
   remains consistent, it also embraces the potential for surprising
@@
 
-Two formal elements from \totan{} are intentionally not included in the
-genre definition above. Perhaps the most clear omission is players' reading
-each others' stories. While we agree that this is a critical element of 
+Two notable formal elements from \totan{} are intentionally excluded from the
+genre definition above. First among these is the oration between players
+of each others' stories. While we agree that this is a critical element of 
 \totan{}---and that it dovetails performatively 
-and practomimetically~\citep{Travis2011} with Scheherazade's
+with Scheherazade's
 performances in the source material---it is not clear that this is
@@
 
-The other formal element omitted from the genre definition is the use of an
+The other formal element excluded from the genre definition is the use of an
 explicit map. The map is an important tangible aspect of \totan{}, 
 in particular with its representation of Baghdad as the largest, most
-important city in the world. However, we also note that virtual spaces
-can be represented without such an explicit map, such as in the MUD family
-of games or choose-your-own-adventure books, both of which
+important city in the world (Figure~\ref{fig:totan-board}). 
+However, we also note that virtual spaces
+can be represented without such an explicit map, such as in text adventures,
+the MUD family of games, or choose-your-own-adventure books, both of which
 use discrete spaces without giving them visual manifestation.
 
-Analyzing \textit{Agents of SMERSH}~\citep{Maxwell2012} establishes that
-these characteristics form a genre and not merely selected elements from one
-game.
+To establish these as characteristics of a genre and not just one game,
+we consider another game that possesses them: 
+\textit{Agents of SMERSH}~\citep{Maxwell2012}.
 \smersh{} is a cooperative game set in the Cold War, with players as 
 covert agents trying to stop the enigmatic Dr.\ Lobo. 
-Players move across the world collecting resources and battling henchmen,
-facing encounters clearly inspired by \totan{}: flip a card to read an
+Players move across the world collecting resources and battling henchmen.
+While the stories are inspired by spy movies and novels,
+the game mechanics are clearly derived from \totan{}: flip a card to read an
 encounter title, choose a reaction from a table, hear an introduction,
-and then get different results based on skills. Unlike \totan{},
-\smersh{} uses dice for story resolution, and supporting resource
-management mechanisms introduce important strategic decisions.
-Despite the use of dice for story resolution, the feedback is still
-primarily narrative.
+and then get different results based on skills.
+Although \smersh{} uses dice for encounter resolution, the feedback is still
+primarily narrative, as illustrated with sample encounters in
+Figure~\ref{fig:encounters}.
 
-For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
-worker-management game about building a city both above and below the
-surface.  \textit{Above and Below} is primarily about building an
-economic engine that generates victory points faster than your
-opponents within a fixed number of rounds. Exploring the underground
-is one source of points, and this is done by determining a random
-entry within its encounter book.  A player reads this entry aloud,
-after which the player is presented with a decision of how to
-react. However, the reaction titles are a thin wrapper around a
-risk-and-reward balance: each has a number which is the number of
-successes required on dice rolls, so players choose the highest number
-they believe they can roll. Higher numbers always equate to better results,
-and this roll is usually the end of the encounter: a successful player
-is given an in-game reward with no other story or narrative.
+For contrast, we also consider \textit{Above and Below}~\citep{Laukat2015}, a
+worker-management game of city-building.  
+%\textit{Above and Below} is primarily about building an
+%economic engine that generates victory points faster than your
+%opponents within a fixed number of rounds,
+Underground exploration includes  player narration
+from an encounter book as in \totan{} and \smersh{}, which at first
+glance may appear to categorize it as a culture-narration game.
+However, in \textit{Above and Below}, reaction choices are thin
+wrappers around risk-and-reward balance:
+an option is essentially a gamble on being able to roll better
+results on the dice, with higher numbers always having better results.
+The feedback is primarily in terms of formal rewards, with few
+encounters providing any narrative feedback, as illustrated in
+the representative encounter of Figure~\ref{fig:encounters}.
 Hence, although \textit{Above and Below} includes an encounter book
@@
 
+\begin{figure}
+\newcommand{\colwidth}{0.48\columnwidth}
+% http://tex.stackexchange.com/questions/2441/how-to-add-a-forced-line-break-inside-a-table-cell
+\newcommand{\specialcell}[2][t]{%
+  \begin{tabular}[#1]{@{}p{\colwidth}@{}}#2\end{tabular}}
+\centering
+\begin{tabular}{|p{\colwidth}|p{\colwidth}|}\hline
+\specialcell{
+\\
+\totan{} encounter \#1965\\
+\\
+Many seaside villages have reported sightings and attacks from
+a large pirate fleet. There are far too many ships in the fleet to attack
+on your own, so you need to find a powerful prince who can stand up to
+the pirate captains.\\
+\\
+\textit{Use Storytelling Skill}\\
+\\
+You present the plight of the villagers with such eloquence and
+style that every prince you address promises to aid in the
+defeat of these rampaging pirates. When you see the size of the
+fleet sent to destroy the renegades, you have no doubt in the eventual
+triumph of justice. [Gain 1 Destiny Point and Courtly Graces skill.]
+}
+&
+\specialcell{
+\\
+\smersh{} encounter \#20.4.1\\
+\\
+You have uncovered evidence that UN Agent F9 also works for the CIA.
+You're not convinced that's necessarily a bad thing, but the Director
+is taking it very seriously. You must find out if he is a threat.\\
+\\
+You meet with F9 and act as if everything is perfectly normal, exchanging
+information about a case you are working together. When he leaves, you follow
+him in your Ferrari.\\
+\\
+\textit{Successfully pass Spycraft test using Driving skill}\\
+\\
+He is highly experienced at spotting tails, but you are equally experienced
+in avoiding being spotted. You follow him to a hotel where he meets
+with a Soviet diplomat. Never mind the CIA: it seems he's working for the
+KGB. [Improve any two skills.]
+}
+\\\hline
+\specialcell{
+\\
+\textit{Above and Below} encounter \#62\\
+\\
+You follow a square tunnel, the walls smooth and masterfully cut.
+Before long you reach a tall chamber; in the center sits a great crystal,
+floating in mid-air. A dim, violet glow emanates from it and as you approach
+a voice fills the air. ``I am the crystal of oracles. If you can prove
+your worth, I will give you knowledge.'' Do you continue speaking to the
+crystal or explore some of the connecting passages?\\
+\\
+\textit{Explore the passages}:\\
+3 successes: gain two coins\\
+5 successes: gain a coin, a fish, and a mushroom
+}
+& 
+\specialcell{
+\\
+\textit{Monster Journal} encounter\\
+\\
+As you're sitting on the edge of a river, you see a large turtle swimming towards you. When it gets close it launches from the water to reveal it's not a turtle, but a kappa!\\
+\\
+\textit{Use Logic skill}\\
+\\
+Remembering Japanese manners, you give the kappa a greeting bow. It returns your bow, emptying the water-filled dip on its head. It loses its strength and crawls to the water. [Gain 1 Inspiration]
+}
+\\\hline
+\end{tabular}
+\caption{Sample encounters}
+\label{fig:encounters}
+\end{figure}
+
 \section{Cultural Empathy}
@@
 We have discussed how \totan{} uses cultural understanding, not only on a
-small scale as a way to reward players, but also as a key takeaway of
+small scale as a way to reward players, but also as a learning outcome of
 the game. In order for the game to reward players for their cultural
-understanding, we have to present cultures in a way that they can, to
-some extent, be understood. In other words, our game needs to foster a
+understanding, cultures must be presented such that they can be understood
+by the player. In other words, the game needs to foster a
 sense of \textit{cultural empathy}, defining ``empathy'' as 
@@
 
-This goal of cultural empathy differs from a goal of \textit{cultural
-literacy} which prepares individuals to participate in another culture
+Cultural empathy is distinct from \textit{cultural
+literacy}, which prepares individuals to participate in another culture
 based on the information of the culture that they have been 
@@
 cultural understanding and their ability to ``perform'' in the given
-culture because of the information they have. This is demonstrated in
+culture because of the information they have---a form of what
+\citet{Travis2011} calls \textit{practomime}.
+This is demonstrated in
 the example above, understanding the cultural views of marriage and
-sex-change have an important role in game play. However, cultural
-empathy makes the main point the understanding, not the
+sex-change have an important role in the game. However, cultural
+empathy makes understanding the main point, not
 performance. \totan{} relies on an understanding of cultural values as
@@
 away from the game they would not have gained the ability to
-participate in either a Middle Eastern or South Asian culture;
-however, they will have gained an understanding of that culture. 
+participate in a Middle Eastern culture,
+although they will have gained some understanding of that culture. 
 %This
@@
 %culture, than about facts about the culture. 
-In the context of the
-game, cultural empathy is a means to the end of cultural
+Hence, from a learning game design perspective,
+cultural empathy is a means to the end of cultural
 literacy. However, outside of the game, focusing on cultural literacy
-within a semi-fictional world works as the means for cultivating
-cultural empathy, particularly world culture empathy.
+within a semi-fictional world is the means for cultivating
+cultural empathy.
 
-\section{Context}
+\section{Context and Theme}
 
-Investigating the potential for culture-narration games to teach
-cultural empathy brought our team into a partnership with
-\textit{[a local library branch hidden for blind review]}. This branch
-serves primarily patrons from low socio-economic status, although
-its special technology education programs also draw homeschoolers from
-around the region.
+In order to explore the potentials within this genre,
+we formed a team to conduct a design experiment and pilot study.
+The team adopted a constructionist perspective~\citep{Papert2001},
+that our understanding of the genre would be improved by engaging
+with design, development, and evaluation---building prototypes
+as a way to understand them.
+We formed a partnership with \textit{[omitted for blind review]}, 
+a non-lending branch that 
+provides community, training, and technology services in a
+low socioeconomic area of our city.
 The youth patrons of the library tend to be 4rd--6th grade 
 ``latchkey kids'' who 
-see the library as a place to visit with friends, to participate in programs,
-and to access technology.
+see the library as an after-school destination to visit with friends, 
+to participate in programs, and to access technology.
 
 Conversations with program coordinators about culture-narration games
-revealed that despite other differences, the youth patrons shared
-a fascination with monsters, horror, and the macabre. 
-This is in part because of cultural taboo: the youth will brag about
-being able to see a PG-13-rated movie or play an M-rated game,
-for example, and several kids worked together to produce an original
-short movie based on battles between movie monsters.
+revealed that despite other differences, the youth patrons shared a
+fascination with monsters, horror, and the macabre.  This answer in
+some ways seemed unsatisfactory: horror and empathy---cultural or
+otherwise---do not seem to be a good match. Indeed,
+to enjoy horror requires an antithetical dehumanization of its subjects.
+However, the love of horror and the monstrous can be
+seen throughout our culture, in all forms of media. Indeed, monsters have
+been a hook for narratives throughout history. They have attracted
+people of various walks of life, by playing into shared human and cultural
+fears.
 
-The fascination exhibited by the youth patrons reflects various theses
-of Cohen's monster theory~\citep{Cohen1996}. 
-We see in their behavior that ``the same creatures who terrify and
-interdict can evoke potent escapist fantasies''~(p.16).
-Recognizing the youth patrons' attraction to this theme,
-Cohen's theory establishing monsters as reflections of culture,
-and the narrative potential of monsters,
-our team decided to engage in a design experiment to validate our
-hypotheses about culture-narration games,
-in collaboration with \textit{the library branch}.
+\citet{Cohen1996} draws attention to ``reading cultures through the
+monsters they engender'' (p.~3).  He notes that the conclusions we
+come to about cultures, from monsters or otherwise, will be a
+combination of many fragments rather than a ``smooth epistemological
+whole.'' A monster is brought about at a particular time and
+represents not only the fears prevalent in the culture but also the
+values, that lead to that fear (p.~4--12). Pinning that down is
+difficult. The monsters work more as a snapshot of the culture at
+a particular moment than a definition. This point is especially
+important for us to note when creating a game whose goal is to
+encourage children's efforts to see something from others'
+perspectives.
+However, Cohen's theory also helps us understand children's fascination
+with monsters, that ``the same creatures who terrify and interdict can evoke potent escapist fantasies''~(p.16).
+
+Our team therefore chose monsters as a key dramatic element for our 
+design experiments. They were not only enticing to our target
+audience---first observed by our community partner and later proven
+through play testing---but Cohen's theory also 
+suggests that they also provide an excellent passage for
+encouraging empathy. Monsters represent
+various cultures' sameness, in base forms of fear, yet difference, in
+the ways that fear becomes specific and embodied. By encouraging an
+understanding of both difference and sameness between cultures, they
+encourage individuals to try to see something from someone else's
+viewpoint.
 
@@
 
-A team of eleven undergraduate students, working with a faculty mentor,
-undertook a design experiment and pilot study to explore a monster-themed
-culture-narration game;
-our approach draws upon constructionist~\citep{Papert2001}
-traditions, engaging in the design experiment with a multifaceted approach,
-building prototypes as a way to understand them.
-The students' majors include Computer Science, English
+Our team comprised ten undergraduate students and one faculty mentor.
+The students' majors included Computer Science, English
 (Creative Writing), Philosophy, Telecommunications (Audio Production),
-and Art (Animation).
-Each student earned three credit hours for the experience during a fifteen-week
+and Art (Animation), and 
+each earned three credit-hours during a fifteen-week
 semester,
@@
 to teach cultural empathy;
-and to use technology to address complications of analog implementations.
+and to use technology to address problems of analog implementations.
 
-Given these goals, the team proceeded with an incremental and
+The team proceeded with an incremental and
 iterative approach to game design and development.
@@
 produced an executable release in accordance with best practices of
-agile software development~\citep[see, for example,][]{Cockburn2006,Keith2010}. 
-Figures~\ref{fig:screenshots} and~\ref{fig:paper-prototype} demonstrate
-design artifacts from the design and development process.
+agile software development~\citep{Cockburn2006};
+figures~\ref{fig:screenshots} and~\ref{fig:paper-prototype} demonstrate
+artifacts from the design and development process.
 Each iteration was tested at
 \textit{[the library branch]}, providing valuable feedback to the
-team.  Each iteration concluded with a retrospective
-meeting~\citep{Kerth2001}, which provided an opportunity for the
-team's reflective practice~\citep{Schon1984}.
-The results of these retrospectives included large and small changes
-to the team's methodology, including a transition from
-story mapping~\citep{Patton2014} to Scrum~\citep{Schwaber2013}.
+team. 
 In addition to paper prototypes, the team 
-developed prototypes using JavaFX, PlayN, and Polymer, 
+developed software-based prototypes using JavaFX, PlayN, and Polymer, 
 all of which allow for HTML5 Web deployment.
@@
 games, monsters, world cultures, and literacy.
+Each iteration concluded with a retrospective
+meeting, which provided an opportunity for the
+team's reflective improvement~\citep{Kerth2001,Schon1984}.
 
-\newcommand{\figwidth}{2.5in}
 \begin{figure}
@@
 
-\section{Pilot Study}
+The title and theme of our game changed between iterations as we came
+to better understanding of the interplay between genre, rules, and our
+players.  The final version---\textit{Monster Journal}---positioned
+two players as rival authors seeking inspiration from monsters around
+the world. The players travel across Africa, Asia, Europe, Oceania,
+and North and South America to encounter monsters from local folklore.
+Encounters are patterned after \totan{} as illustrated in
+Figure~\ref{fig:encounters}, and the game exhibits the five characteristics
+identified for cultural-narration games.
 
-\textit{In which we describe what playtesting has revealed to us}
+While the visual design and overall rules structure remained stable
+throughout the iterations of our experiment, the encounter stories
+themselves changed dramatically.
+Our target audience required that both stories and gameplay 
+had to be much shorter than in other games that we analyzed,
+our final version having a 175-character limit.
+\textit{[Library branch]} is a high-distraction environment,
+where our game competes for attention against both formal and informal
+activities. The game takes approximately fifteen minutes to complete,
+which we found to be a good balance of keeping players' attention,
+making them curious to play again,
+and fitting within our limited means for producing content.
 
@@
 cultural empathy.
+These games combine the educational value of learning from stories
+and learning from games, and they provide hooks for designers to 
+emphasize various positive skills, including literacy, 
+oratory, cultural empathy, history, mathematics, and strategic thinking.
 However, as with any educational intervention, these games do not
 come without complications.
+As others have pointed out~\citep[\textit{e.g.}]{Slota2015}, 
+the exact role of narrative in games and in learning is not well
+understood. Hence, we focus our conclusions on more immediate
+goals of continued fruitful experimentation with culture-narration games.
 
@@
 program. With curriculum aides and improved scaffolding, such as
-post-game debriefing, we are certain the game could be easily
+post-game debriefing~\citep{Nicholson2012}
+or reflections~\citep{Hickey2013},
+we are certain the game could be easily
 integrated into school environments, particularly those with

### 2016-03-24 paper.tex
Remove TotAN board image to reduce page count to 7

@@
 \newcommand{\smersh}{\textit{SMERSH}}
+\newcommand{\figwidth}{2.5in}
 
@@
 Story and Destiny point goal while having a successful encounter
-in Baghdad---the City of Peace. Notice in Figure~\ref{fig:totan-board}
-that Baghdad is the center of the world, much larger and more beautiful
-than any other city.
-
-\newcommand{\figwidth}{2.5in}
-%\begin{wrapfigure}{R}{\figwidth}
-\begin{figure}[b]
-\centering
-\includegraphics[width=\figwidth]{totan-board}
-\caption{Center of the \textit{Tales of the Arabian Nights} game board~\citep{Goldberg2009}.}
-\label{fig:totan-board}
-%\end{wrapfigure}
-\end{figure}
+in Baghdad---the City of Peace.
 
@@
 in particular with its representation of Baghdad as the largest, most
-important city in the world (Figure~\ref{fig:totan-board}). 
+important city in the world. 
 However, we also note that virtual spaces

### 2016-03-24 paper.tex
Revise abstract to match the body

@@
 \begin{adjustwidth}{0.5in}{0.5in}
-  \textbf{Abstract:}  \textit{Culture-narration games} are a promising
-  genre for transformative games. This genre can be defined formal and
-  dramatic characteristics, and we identify tabletop games that belong
-  to the genre.  Adding technological components to the game affords
-  new interaction modes as well as convenient digital distribution. We
-  describe our team's pilot project at creating a technology-enhanced
-  culture-narration game to improve cultural empathy among players aged
-  10--14.  This pilot project highlights several complications with
-  this genre, including thematic, technical, and production
-  challenges.  However, the pilot project justifies continued
-  experimentation, development, and formal assessment within the
-  genre.
+  \textbf{Abstract:} We are inspired by the educational potential of
+  the board game \textit{Tales of the Arabian Nights}, which is a
+  popular game among enthusiasts that teaches the player about the
+  eponymous folk tale and the culture around it. Consider this game
+  and others like it, we identify the characteristics that define a
+  genre of \textit{culture-narration games}, which we consider to have
+  untapped potential for educational and transformative games.
+  We describe a design experiment through which a multidisciplinary team
+  followed an iterative and incremental process, in collaboration
+  with a community partner, to investigate the potential of this genre.
+  The result is a game that uses a theme of monsters from around the
+  world to teach cultural empathy.
+  This pilot project reveals both the promise and the several complications
+  with the genre, which lead to recommendations for future work.
 \end{adjustwidth}

### 2016-03-25 paper.tex
Update game name
Update paper title

@@
 
-\newcommand{\projecttitle}{The Culture-Narration Genre: Defining the genre and identifying its potential for educational games}
+\newcommand{\projecttitle}{Culture-narration games: a definition and pilot study}
 
@@
 \\
-\textit{Monster Journal} encounter\\
+\textit{Traveler's Notebook} encounter\\
 \\
@@
 to better understanding of the interplay between genre, rules, and our
-players.  The final version---\textit{Monster Journal}---positioned
-two players as rival authors seeking inspiration from monsters around
-the world. The players travel across Africa, Asia, Europe, Oceania,
-and North and South America to encounter monsters from local folklore.
-Encounters are patterned after \totan{} as illustrated in
-Figure~\ref{fig:encounters}, and the game exhibits the five characteristics
-identified for cultural-narration games.
+players.  The final version---\textit{Traveler's Story: Monster
+  Tales}---positioned two players as rival authors seeking inspiration
+from monsters around the world. The players travel across Africa,
+Asia, Europe, Oceania, and North and South America to encounter
+monsters from local folklore.  Encounters are patterned after \totan{}
+as illustrated in Figure~\ref{fig:encounters}, and the game exhibits
+the five characteristics identified for cultural-narration games.
 

### 2016-03-25 paper.tex
Incorporated Ash's edits

@@
   \textbf{Abstract:} We are inspired by the educational potential of
-  the board game \textit{Tales of the Arabian Nights}, which is a
-  popular game among enthusiasts that teaches the player about the
-  eponymous folk tale and the culture around it. Consider this game
+  the board game \textit{Tales of the Arabian Nights}, a
+  popular game based on the eponymous folk tale. Considering this game
   and others like it, we identify the characteristics that define a
@@
   world to teach cultural empathy.
-  This pilot project reveals both the promise and the several complications
+  This pilot project reveals both the promise and several complications
   with the genre, which lead to recommendations for future work.
@@
 \textit{Tales of the Arabian Nights}~\citep{Goldberg2009} is a 
-tabletop board game based on the eponymous folk tale,
-originally published in 1985~\citep{Goldberg1985} and
-released with many revisions in 2009.
-It is popular among board game hobbyists, having a board game
-rank of~239 (thematic rank~75) on Board Game Geek as of this writing,
+tabletop board game based on the eponymous folk tale. It was
+originally published in 1985~\citep{Goldberg1985}, and it was
+re-released with many revisions in 2009.
+It is popular among board game hobbyists:
+on Board Game Geek, it has a rank of~239 (thematic rank~75),
 and Shut Up \& Sit Down rank it as the ninth best game of all 
@@
 setting, exploring the known world to accumulate Story and
-Destiny points. The winner is the one who is able to meet their 
+Destiny points. The winner is the player who is able to meet their 
 Story and Destiny point goal while having a successful encounter
@@
 such as the Sindbad's escape from the valley of diamonds or Aladdin's
-being trapped in a magic cave.
+entrapment in a magic cave.
 
@@
 particularly  the youth-friendly translation by \citep{Neil1994},
-which made us recognize the cleverness of the design
+which revealed to us the cleverness of the design
 and its elegant dovetailing with the source material.
 Indeed, although the game was not explicitly designed as a learning game,
-we find in it exhibits many of the learning game design principles
+we find it exhibits many of the learning game design principles
 established by \citet{Klopfer2009}.
@@
 
-\section{Defining the genre}
+\section{Defining the Genre}
 
@@
   remain unexpected twists: perhaps fighting the Powerful Ifrit without
-  combat skills makes him respect you and grant a boon.  \totan{},
-  after all, is a game that allows the player to choose to ``drink'' a
-  Violent Storm or ``enter'' a Mysterious Artifact.  While the world
+  combat skills makes him respect you and grant a boon. 
+  While the world
   remains consistent, it also embraces the potential for surprising
@@
 \smersh{} is a cooperative game set in the Cold War, with players as 
-covert agents trying to stop the enigmatic Dr.\ Lobo. 
+covert agents trying to stop the entigmatic Dr.\ Lobo---a mad scientist
+bent on world domination.
 Players move across the world collecting resources and battling henchmen.
@@
 that players read to each other, it does not satisfy the characteristics
-to be in the culture-narration game genre.
+of the culture-narration game genre.
 
@@
 based on the information of the culture that they have been 
-given~\citep[p.165]{Hirsch1983}. Being culturally literate is having the
-ability to be able to participate in another culture that you are
+given~\citep[p.165]{Hirsch1983}. Being culturally literate means having the
+ability to participate in another culture that you are
 ``dropped into,'' similar to how being literate is having the ability to
@@
 see the library as an after-school destination to visit with friends, 
-to participate in programs, and to access technology.
+participate in programs, and access technology.
 
@@
 seen throughout our culture, in all forms of media. Indeed, monsters have
-been a hook for narratives throughout history. They have attracted
-people of various walks of life, by playing into shared human and cultural
-fears.
+been a hook for narratives throughout history.
 
@@
 suggests that they also provide an excellent passage for
-encouraging empathy. Monsters represent
+developing empathy. Monsters represent
 various cultures' sameness, in base forms of fear, yet difference, in
-the ways that fear becomes specific and embodied. By encouraging an
+the ways that fear becomes specific and embodied. By engendering an
 understanding of both difference and sameness between cultures, they
@@
 The title and theme of our game changed between iterations as we came
-to better understanding of the interplay between genre, rules, and our
+to a better understanding of the interplay between genre, rules, and our
 players.  The final version---\textit{Traveler's Story: Monster

### 2016-06-06 paper.tex
Add second paper prototype image
There was some ugly whitespace, and adding another image helped achieve balance.
Add screenshots from iteration 3
Remove blinded elements of accepted paper

@@
 \newtoggle{blind}
-\toggletrue{blind}
+\togglefalse{blind}
 
@@
 as a way to understand them.
-We formed a partnership with \textit{[omitted for blind review]}, 
-a non-lending branch that 
+We formed a partnership with Connection Corner, 
+a non-lending branch of Muncie Public Library~(Muncie, Indiana, USA) that 
 provides community, training, and technology services in a
 low socioeconomic area of our city.
-The youth patrons of the library tend to be 4rd--6th grade 
+Many youth patrons of the library are 4rd--6th grade 
 ``latchkey kids'' who 
@@
 semester,
-working together in a studio space for two hours each Monday, Wednesday,
-and Friday.
-% In future version, cite the TOCE paper
+working together in an academic studio~\citep{Gestwicki2016}
+for two hours each Monday, Wednesday, and Friday.
 The team collectively agreed upon three primary goals for the design
@@
 Each iteration was tested at
-\textit{[the library branch]}, providing valuable feedback to the
+Connection Corner, providing valuable feedback to the
 team. 
@@
 \includegraphics[width=\figwidth]{paper-prototype}
-\caption{Paper prototype from Iteration 3}
+\includegraphics[width=\figwidth]{paper-prototype-2}
+\caption{Paper prototypes from Iteration 3}
+\label{fig:paper-prototype}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\figwidth]{i3-shot1}
+\includegraphics[width=\figwidth]{i3-shot2}
+\caption{Screenshots from Iteration 3}
 \label{fig:paper-prototype}
@@
 our final version having a 175-character limit.
-\textit{[Library branch]} is a high-distraction environment,
+Connection Corner is a high-distraction environment,
 where our game competes for attention against both formal and informal
@@
 \section{Acknowledgments}
-\textit{The team, organizational, and community partner
-  acknowledgments are hidden for blind review.}
+The authors acknowledge the contributions of other team members who
+created \textit{Traveler's Notebook}: 
+Nicholas Boyer,
+Ashley Cox,
+Jessica Lohse,
+Jack McGinnis,
+Lucas Miller,
+Joshua Schoen,
+Brennan Wade,
+and Caleb Woods.
+Jennifer Grouling and Joyce Huff also contributed to our design experiment.
+We are grateful for the cooperation of Muncie Public Library, especially
+from Rebecca Parker and Stuart Cotton 
+This work was internally supported by the College of Sciences and
+Humanities and the Computer Science Department at Ball State University.
 

### 2016-06-07 paper.tex
Tighten prose to get down to seven pages

@@
 
-We propose that \totan{} is one of very few published games comprising a genre
-of \textit{culture-narration games}. Recognizing that the term ``genre''
-can be interpreted in multiple ways, we adopt the perspective established
-by \citet{Arsenault2009},
-that a genre is ``the codified usage of particular mechanics
-and game design patterns to express a range of intended play-experiences.''
-We define this genre, then, as having the following characteristics.
+We propose that \totan{} is one of very few published games comprising
+a genre of \textit{culture-narration games}.  This follows the
+definition for ``genre'' given by \citet{Arsenault2009}---that it is
+``the codified usage of particular mechanics and game design patterns
+to express a range of intended play-experiences.''  We define
+\textit{culture-narration games}, then, as having the following
+characteristics:
 
@@
 \item \textit{Reward players for decisions that reflects
-    cultural understanding.}  The knowledge that a player brings to
-  bear on decisions is not just knowledge of in-game systems. 
-  A player can predict that choosing to fight a Powerful Ifrit 
- without the Weapon Use skill will not end well,
-  but this prediction is not made by having memorized the 
-  \textit{Book of Tales}.
+    cultural understanding.}  The knowledge that a player uses in the game
+  is not just knowledge of in-game systems. 
+  For example, it is not by memorizing the \textit{Book of Tales}
+  that leads a player to predict that fighting a Powerful Ifrit without
+ the Weapon Use skill will not end well.
   Rather, this prediction draws upon a cultural understanding of power,
   spirits, violence, and much more.
-  Note that the game rewards player decisions without resorting to
-  clear-cut ``right'' and ``wrong'' (``moral/immoral'',
+  The game rewards player decisions without resorting to
+  binary``right'' and ``wrong'' (``moral/immoral'',
   ``light side/dark side'', ``paragon/renegade'') choices.  There
@@
   While the world
-  remains consistent, it also embraces the potential for surprising
+  remains consistent, it also has the potential for surprising
   results.
@@
 
-For contrast, we also consider \textit{Above and Below}~\citep{Laukat2015}, a
+For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
 worker-management game of city-building.  
@@
 Underground exploration includes player narration
-from an encounter book as in \totan{} and \smersh{}, which at first
+from an encounter book, which at first
 glance may appear to categorize it as a culture-narration game.
 However, in \textit{Above and Below}, reaction choices are thin
-wrappers around risk-and-reward balance:
+wrappers around probabilistic risk:
 an option is essentially a gamble on being able to roll better
@@
 
-We have discussed how \totan{} uses cultural understanding, not only on a
+\totan{} uses cultural understanding, not only on a
 small scale as a way to reward players, but also as a learning outcome of
@@
 through play testing---but Cohen's theory also 
-suggests that they also provide an excellent passage for
+suggests that they provide an affordance for
 developing empathy. Monsters represent
@@
 
-Physical games such as \totan{} and \smersh{} are expensive to produce
-and, for many, intimidating to encounter. \totan{}'s 19-page
-instruction booklet may not faze a board game enthusiast, but we have
-observed that others are intimidated by the large rulebook, heavy box,
-and many components included in the game. Combined with the duration
-of play and complexity of rules, such games seem inappropriate for use
+\totan{}'s 19-page
+instruction booklet may not faze a games enthusiast, but 
+many are intimidated by the large rulebook, heavy box, and 
+number of components.
+Combined with the duration
+of play, complexity of rules, and relatively high price,
+such games are inappropriate for use
 in almost all formal educational settings such as conventional
-schools.  However, we have found that the combination of technology
+schools.  However, the combination of technology
 enhancement, grade-level writing, and shorter duration circumvents
